---
title: "S&P 500 Backtesting Comparison"
author: "Hansen Han"
date: "2023-06-13"
output: html_document
---

We are taking the output of 100 runs w/ Simple Moving Average (SMA) crossover and Mean Reversion w/ and without shorting to see which strategy is better and what parameters are associated with higher returns when trading an S&P 500 index fund (in this case SPY)

## Load Libraries
```{r}
library(tidyverse)
library(magrittr)
```

## Load Data
```{r}
df1 <- read_csv("run_output/mean_reversion_no_shorting_100_runs.csv")
df1$strategy <- "Mean Reversion"
df1$shorting <- FALSE

df2 <- read_csv("run_output/mean_reversion_shorting_allowed_100_runs.csv")
df2$strategy <- "Mean Reversion"
df2$shorting <- TRUE

df3 <- read_csv("run_output/sma_crossover_no_shorting_100_runs.csv")
df3$strategy <- "SMA Crossover"
df3$shorting <- FALSE

df4 <- read_csv("run_output/sma_crossover_shorting_allowed_100_runs.csv")
df4$strategy <- "SMA Crossover"
df4$shorting <- TRUE


#combine all datasets
all_data <- df1 %>% full_join(df2) %>% full_join(df3) %>% full_join(df4)
all_data
```

Create some labels for the 4 types of runs being compared (Mean Reversion Shorting, Mean Reversion Regular, SMA Crossover Shorting, Mean Reversion Regular)
```{r}
all_data %<>% mutate(shorting_label = case_when(
  shorting %in% c(FALSE) ~ "Regular",
  shorting %in% c(TRUE) ~ "Shorting"
)) %>% mutate(run_type = paste0(strategy, "_", shorting_label)) %>% select(-shorting_label)
```

Drop columns w/o data
```{r}
all_data %<>% select(-`...1`, -profitable_quarters, -total_trades)
```

## Analysis


### How do the different strategies compare?

Regular Scale
```{r}
p1 <- all_data %>% mutate(shorting = case_when(
    shorting == FALSE ~ "Regular",
    shorting == TRUE ~ "Shorting Allowed"
  )) %>% ggplot(aes(x = shorting, y = cumulative_return*100, color=shorting)) + geom_boxplot(outlier.shape = NA) + geom_jitter(size=0.5) + theme_bw() + xlab("") + ylab("Cumulative Return %") + theme(legend.position = "none") +
  geom_hline(yintercept = 2.07, color = "red", size = 0.5, alpha = 0.5) + facet_grid(~ strategy) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") 
p1 
```

Log Scale
```{r}
p2 <- all_data %>% mutate(shorting = case_when(
    shorting == FALSE ~ "Regular",
    shorting == TRUE ~ "Shorting Allowed"
  )) %>% ggplot(aes(x = shorting, y = cumulative_return*100, color=shorting)) + geom_boxplot(outlier.shape = NA) + geom_jitter(size=0.5) + theme_bw() + xlab("") + ylab("Cumulative Return % (log10)") + theme(legend.position = "none") + scale_y_log10() + coord_cartesian(ylim = c(0.01, 100)) +
  geom_hline(yintercept = 2.07, color = "red", size = 1, alpha = 0.5) + facet_grid(~ strategy) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
p2
```

What percent of each run beat buy & hold? 
```{r}
count_greater_than_threshold <- all_data %>%
  group_by(strategy, shorting) %>%
  summarise(count = sum(cumulative_return > 0.02075725)) %>% mutate(shorting = case_when(
    shorting == FALSE ~ "Regular",
    shorting == TRUE ~ "Shorting Allowed"
  )) 


p3 <- ggplot(count_greater_than_threshold, aes(x = shorting, y = count)) +
  geom_bar(stat = "identity", alpha = 0.6, aes(fill = ifelse(count > 50, "green", "red"))) +
  labs(x = "", y = "% of Runs that Beat Buy & Hold", title = "") +
  coord_cartesian(ylim = c(0, 100)) +
  theme_bw() +
  geom_hline(yintercept = 50, color = "grey", size = 1, alpha = 0.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  scale_fill_manual(values = c("red" = "red", "green" = "green")) + facet_grid(~ strategy)

p3
```

### What factors affect performance? 

Mark data as outperforming or underperforming the market...
```{r}
all_data$greater_than_threshold <- ifelse(all_data$cumulative_return > 0.02075725, 1, 0)
```

#### Mean Reversion

```{r}
plot_function <- function(all_data, y_col, y_label, scale_log = FALSE) {
  library(ggpubr)
  
  data_processed <- all_data %>%
    filter(strategy == "Mean Reversion") %>%
    mutate(
      run_label = case_when(
        greater_than_threshold == 0 ~ "Underperforms",
        greater_than_threshold == 1 ~ "Outperforms"
      ),
      shorting = case_when(
        shorting == TRUE ~ "Shorting Allowed",
        shorting == FALSE ~ "Regular"
      )
    )
  
  plot <- data_processed %>%
    ggplot(aes(x = run_label, y = !!sym(y_col), color = run_label)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(size = 0.5) +
    theme_bw() +
    xlab("Performance vs. Buy & Hold") +
    ylab(y_label) +
    theme(legend.position = "none") +
    ggtitle(y_label) +
    stat_compare_means(comparisons = list(c("Underperforms", "Outperforms"))) +
    facet_grid(~ shorting)
  
  if (scale_log) {
    plot <- plot + scale_y_log10()
  }
  
  return(plot)
}

# Now you can use this function for any column in your dataset

p4 <- plot_function(all_data, "buy_threshold", "Buy Threshold")
p5 <- plot_function(all_data, "take_profit_threshold", "Take Profit Threshold")
p6 <- plot_function(all_data, "stop_loss_threshold", "Stop Loss Threshold")
p7 <- plot_function(all_data, "ma", "Moving Average Length (hours)")

p4
p5
p6
p7
```

#### SMA Crossover

```{r}
plot_function <- function(all_data, y_col, y_label, scale_log = FALSE) {
  library(ggpubr)
  
  data_processed <- all_data %>%
    filter(strategy == "SMA Crossover") %>%
    mutate(
      run_label = case_when(
        greater_than_threshold == 0 ~ "Underperforms",
        greater_than_threshold == 1 ~ "Outperforms"
      ),
      shorting = case_when(
        shorting == TRUE ~ "Shorting Allowed",
        shorting == FALSE ~ "Regular"
      )
    )
  
  plot <- data_processed %>%
    ggplot(aes(x = run_label, y = !!sym(y_col), color = run_label)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(size = 0.5) +
    theme_bw() +
    xlab("Performance vs. Buy & Hold") +
    ylab(y_label) +
    theme(legend.position = "none") +
    ggtitle(y_label) +
    stat_compare_means(comparisons = list(c("Underperforms", "Outperforms"))) +
    facet_grid(~ shorting)
  
  if (scale_log) {
    plot <- plot + scale_y_log10()
  }
  
  return(plot)
}

# Now you can use this function for any column in your dataset

p8 <- plot_function(all_data, "ma1_length", "Short Moving Average Length (MA1)")
p9 <- plot_function(all_data, "ma2_length", "Long Moving Average Length (MA2)")

p8
p9
```