---
title: "Cryptocurrency Strategy Comparison"
author: "Hansen Han"
date: "2023-06-08"
output: html_document
---

The goal of this analysis is to compare mean reversion and trend following strategies that have shorting/no shorting and fees/no fees to see what factors affect returns and see if any are viable for potential trading.

## Load Libraries
```{r}
library(tidyverse)
library(magrittr)
```

## Load Data
```{r}
df1 <- read_csv("run_output/mean_reversion_no_shorting_no_fee.csv")
df1$strategy <- "Mean Reversion"
df1$shorting <- FALSE
df1$fees <- FALSE

df2 <- read_csv("run_output/mean_reversion_no_shorting_with_fee.csv")
df2$strategy <- "Mean Reversion"
df2$shorting <- FALSE
df2$fees <- TRUE

df3 <- read_csv("run_output/mean_reversion_shorting_no_fee.csv")
df3$strategy <- "Mean Reversion"
df3$shorting <- TRUE
df3$fees <- FALSE

df4 <- read_csv("run_output/mean_reversion_shorting_with_fee.csv")
df4$strategy <- "Mean Reversion"
df4$shorting <- TRUE
df4$fees <- TRUE

df5 <- read_csv("run_output/sma_crossover_no_shorting_no_fee.csv")
df5$strategy <- "SMA Crossover"
df5$shorting <- FALSE
df5$fees <- FALSE

df6 <- read_csv("run_output/sma_crossover_no_shorting_with_fee.csv")
df6$strategy <- "SMA Crossover"
df6$shorting <- FALSE
df6$fees <- TRUE

df7 <- read_csv("run_output/sma_crossover_shorting_no_fee.csv")
df7$strategy <- "SMA Crossover"
df7$shorting <- TRUE
df7$fees <- FALSE

df8 <- read_csv("run_output/sma_crossover_shorting_with_fee.csv")
df8$strategy <- "SMA Crossover"
df8$shorting <- TRUE
df8$fees <- TRUE

#combine all datasets
all_data <- df1 %>% full_join(df2) %>% full_join(df3) %>% full_join(df4) %>% full_join(df5) %>% full_join(df6) %>% full_join(df7) %>% full_join(df8)
all_data
```

Create some labels for the run types
```{r}
all_data %<>% mutate(shorting_label = case_when(
  shorting %in% c(FALSE) ~ "Regular",
  shorting %in% c(TRUE) ~ "Shorting"
)) %>% mutate(fee_label = case_when(
  fees %in% c(FALSE) ~ "No-Fee",
  fees %in% c(TRUE) ~ "Fees"
)) %>% mutate(run_type = paste0(strategy, "_", shorting_label, "_", fee_label)) %>% select(-shorting_label, -fee_label)

all_data
```

Drop columns w/o data
```{r}
all_data %<>% select(-`...1`, -stdev_return_rate, -sharpe_ratio)
```


## Analysis


### How do the differert strategies compare 
```{r}
#log scale
p1 <- all_data %>% mutate(shorting = case_when(
    shorting == FALSE ~ "Regular",
    shorting == TRUE ~ "Shorting Allowed"
  )) %>% mutate(fees = case_when(
    fees == FALSE ~ "No Fees",
    fees == TRUE ~ "Volume-Based Fees"
  )) %>% ggplot(aes(x = shorting, y = cumulative_return*100, color=shorting)) + geom_boxplot(outlier.shape = NA) + geom_jitter(size=0.5) + theme_bw() + xlab("") + ylab("Cumulative Return (%)") + theme(legend.position = "none") + scale_y_log10() + coord_cartesian(ylim = c(0.01, 100000)) +
  geom_hline(yintercept = 3224.857, color = "red", size = 1, alpha = 0.5) + facet_grid(~ strategy + fees) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")
p1 


```

What percent of each run beat buy & hold? 
```{r}
count_greater_than_threshold <- all_data %>%
  group_by(run_type, strategy, shorting, fees) %>%
  summarise(count = sum(cumulative_return > 32.24857)) %>% mutate(shorting = case_when(
    shorting == FALSE ~ "Regular",
    shorting == TRUE ~ "Shorting Allowed"
  )) %>% mutate(fees = case_when(
    fees == FALSE ~ "No Fees",
    fees == TRUE ~ "Volume-Based Fees"
  ))


p2 <- ggplot(count_greater_than_threshold, aes(x = shorting, y = count)) +
  geom_bar(stat = "identity", alpha = 0.6, aes(fill = ifelse(count > 50, "green", "red"))) +
  labs(x = "", y = "% of Runs that Beat Buy & Hold", title = "") +
  coord_cartesian(ylim = c(0, 100)) +
  theme_bw() +
  geom_hline(yintercept = 50, color = "grey", size = 1, alpha = 0.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  scale_fill_manual(values = c("red" = "red", "green" = "green")) + facet_grid(~ strategy + fees)

p2
```


## Run a regression model for SMA crossover to see what the relationship is between the features and returns

### With Fees
```{r}
sma_crossover_regression_data <- all_data %>% filter(strategy %in% c("SMA Crossover")) %>% filter(fees == TRUE)

# Perform regression analysis
model <- lm(cumulative_return ~ total_trades + shorting, data = sma_crossover_regression_data)

# Display the summary of the regression model
summary(model)
```
### W/o Fees

```{r}
sma_crossover_regression_data <- all_data %>% filter(strategy %in% c("SMA Crossover")) %>% filter(fees == FALSE)

# Perform regression analysis
model <- lm(cumulative_return ~ total_trades + shorting, data = sma_crossover_regression_data)

# Display the summary of the regression model
summary(model)
```

### Consistently profitable strategies that also outperform B&H that are tradable in NY (no shorting, has fees)?

There are 35 quarters, so the highest strategy which was profitable 22/35 quarters is about 62% profitable. 
```{r}
all_data$greater_than_threshold <- ifelse(all_data$cumulative_return > 32.24857, 1, 0)

tradeable_winning_strategies <- all_data %>% arrange(desc(profitable_quarters)) %>% filter(greater_than_threshold == 1) %>% filter(fees == TRUE) %>% filter(shorting == FALSE) %>% arrange(desc(profitable_quarters))

tradeable_losing_strategies <- all_data %>% arrange(desc(profitable_quarters)) %>% filter(greater_than_threshold == 0) %>% filter(fees == TRUE) %>% filter(shorting == FALSE)

```


### Compare them to one another across different parameters


#### Mean Reversion
```{r}
plot_function <- function(all_data, y_col, y_label, scale_log = FALSE) {
  library(ggpubr)
  
  data_processed <- all_data %>%
    filter(strategy == "Mean Reversion", fees == TRUE) %>%
    mutate(
      run_label = case_when(
        greater_than_threshold == 0 ~ "Underperforms",
        greater_than_threshold == 1 ~ "Outperforms"
      ),
      shorting = case_when(
        shorting == TRUE ~ "Shorting Allowed",
        shorting == FALSE ~ "Regular"
      )
    )
  
  plot <- data_processed %>%
    ggplot(aes(x = run_label, y = !!sym(y_col), color = run_label)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(size = 0.5) +
    theme_bw() +
    xlab("Performance vs. Buy & Hold") +
    ylab(y_label) +
    theme(legend.position = "none") +
    ggtitle(y_label) +
    stat_compare_means(comparisons = list(c("Underperforms", "Outperforms"))) +
    facet_grid(~ shorting)
  
  if (scale_log) {
    plot <- plot + scale_y_log10()
  }
  
  return(plot)
}

# Now you can use this function for any column in your dataset

p3 <- plot_function(all_data, "buy_threshold", "Buy Threshold")
p4 <- plot_function(all_data, "take_profit_threshold", "Take Profit Threshold")
p5 <- plot_function(all_data, "stop_loss_threshold", "Stop Loss Threshold")
p6 <- plot_function(all_data, "ma", "Moving Average Length (hours)")
p7 <- plot_function(all_data, "mean_trades_per_quarter", "Median Trades Per Quarter", scale_log = TRUE)
p8 <- plot_function(all_data, "total_trades", "Total Trades", scale_log = TRUE)

p3
p4
p5
p6
p7
p8
```

```{r}
library(patchwork)
# Combine the plots using patchwork
combined_plot <- p3 + p5 + p6 + p7 + p8

# Display the combined plot
combined_plot
```


#### SMA Crossover
```{r}
plot_function <- function(all_data, y_col, y_label, scale_log = FALSE) {
  library(ggpubr)
  
  data_processed <- all_data %>%
    filter(strategy == "SMA Crossover", fees == TRUE) %>%
    mutate(
      run_label = case_when(
        greater_than_threshold == 0 ~ "Underperforms",
        greater_than_threshold == 1 ~ "Outperforms"
      ),
      shorting = case_when(
        shorting == TRUE ~ "Shorting Allowed",
        shorting == FALSE ~ "Regular"
      )
    )
  
  plot <- data_processed %>%
    ggplot(aes(x = run_label, y = !!sym(y_col), color = run_label)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(size = 0.5) +
    theme_bw() +
    xlab("Performance vs. Buy & Hold") +
    ylab(y_label) +
    theme(legend.position = "none") +
    ggtitle(y_label) +
    stat_compare_means(comparisons = list(c("Underperforms", "Outperforms"))) +
    facet_grid(~ shorting)
  
  if (scale_log) {
    plot <- plot + scale_y_log10()
  }
  
  return(plot)
}

# Now you can use this function for any column in your dataset

p9 <- plot_function(all_data, "ma1_length", "Short Moving Average Length (MA1)")
p10 <- plot_function(all_data, "ma2_length", "Long Moving Average Length (MA2)")
p11 <- plot_function(all_data, "mean_trades_per_quarter", "Median Trades Per Quarter", scale_log = TRUE)
p12 <- plot_function(all_data, "total_trades", "Total Trades", scale_log = TRUE)

p9
p10
p11
p12
```


```{r}
library(patchwork)
# Combine the plots using patchwork
combined_plot <- p9 + p10 + p11 +p12

# Display the combined plot
combined_plot
```

